Determinism, Complexity, and Predictability in Computer Performance

### 2014-02-14 pecomplexity.tex
editing

>>>>

We will start with the lowest complexity program: \col. Conceptually this program is very simple but due to computer design choices this program can result in highly complex even chaotic time series traces, see \cite{toddchaos}. If we forecast this signal using an out-of-the-box linear forecaster (ARIMA) or an out-of-the-box constant predictor we receive MASE ($0.59 \pm 0.21$ and $0.57 \pm$ 0.001 respectively) which suggest that on average we can only predict this signal twice as well as the simple random walk prediction. This would suggest that there is some predictive structure present but not that much. However, if we quantify the predictive structure present in this signal using WPE we see that this signal is highly structured with a WPE of ($0.51 \pm 0.003$.) The existence of structured which we quantified with WPE implies that an ideal predictor capable of using this signals information should be able to predict this signal much better than the random walk due to the high amount of predictive strucutre or equivalently the low complexity. This is in fact the case. \col is a complex signal, i.e., chaotic, but it is complex in a structured way. Using a prediction technique that can process nonlinear information yields forecasts with a MASE of $0.05 \pm 0.001$, this is a prediction 20 times more accurate than a random walk forecast. The existence of predictve structure as quantified by WPE could be precisely the information needed to continue searching for an ideal predictor like LMA.

Maybe even more conniving than a low complexity signal like \col is a moderately complex signal like \svd$_5$. This signal omits a WPE of $0.677 \pm 0.006$, this is a significantly higher complexity than \col but still implies the exsitence of a great deal of predictive structure. However, this increase in complexity is enough to really throw off the constant and linear prediction strategies. The MASE scores for ARIMA and na\"ive imply 2.5 and 20 times worse prediction than using a simple random walk forecast. This may be discouragin to someone using an out of the box predictor that the signal is too complex to predict. However by \emph{knowing} that there exists predictable strucutre in this signal we can try to use a more advanced technique, in particular LMA. If we do this we get forecasts which score 1.5 times better than the random walk. Showing that this usable structure (althought hidden to the linear and constant method) was quantifiable by WPE.

Finally, we consider the most complex program: \gcc.
